---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---


## Loading and preprocessing the data

```{r libraries and setup}
library(tidyverse)
library(lubridate)
library(knitr)

theme_set(theme_minimal()) #Change the setup for ggplot graphics.

knitr::opts_chunk$set(echo = TRUE,
                      dev.args = list(png = list(type = "cairo"))) #Added echo so all code is visible, and dev.args for prettier graphics.
```

```{r load and preprocessing data}

unzip("C:/Users/bsmor/OneDrive/Dokumenter/Reproducible_research/RepData_PeerAssessment1-master/activity.zip") 

dt <- read_csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
dt %>% 
  group_by(date) %>% 
  summarise("Total steps" = sum(steps))
```

### histogram of the total amount of steps.

```{r}
dt %>% 
  group_by(date) %>% 
  summarise("total_steps"= sum(steps)) %>% 
  ggplot(aes(total_steps)) +
  geom_histogram()
```



In order to calculate the mean and median, we have to remove the missing values.

```{r}
dt %>% 
  group_by(date) %>% 
  summarise("total_steps" = sum(steps)) %>% 
  na.omit() %>% 
  summarise(mean(total_steps), median(total_steps))
```



## What is the average daily activity pattern?

A line diagram showing the mean of the intervals.
```{r}
dt %>% 
  group_by(interval) %>% 
  na.omit() %>% 
  summarise(Mean = mean(steps)) %>% 
  ggplot(aes(interval, Mean)) +
  geom_line()
```


Interval 835 has the highest mean.
```{r}
dt %>% 
  group_by(interval) %>% 
  na.omit() %>% 
  summarise(mean = mean(steps)) %>%
  filter(mean == max(mean))
```



## Imputing missing values


There are 2304 missing values.

```{r}
sum(is.na(dt))
```


Here we replace the missing values with mean based on which interval the missing value is in. Futhermore we load the new data into the frame "no_na_dt".


```{r}
no_na_dt <- dt %>% 
  group_by(interval) %>% 
  mutate(steps = replace_na(steps, mean(steps, na.rm = TRUE))) %>% 
  ungroup()
```



Histogram of the total number of steps taken each day

```{r}
no_na_dt %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps)) %>% 
  ggplot(aes(total_steps)) +
  geom_histogram()
```

The mean and median total number of steps taken per day. The values are close to the original values, but there is a difference. 

```{r}
no_na_dt %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps)) %>% 
  summarise(mean(total_steps), median(total_steps))
```


## Are there differences in activity patterns between weekdays and weekends?

Yes, there are more steps in the earlier intervals in the weekdays. However across all intervals the weekend has more steps.


```{r}
no_na_dt %>% 
  mutate(weekday = wday(date, label =F),
         weekday = case_when(weekday >= 2 & weekday <= 6 ~ 1,
                             weekday == 1 | weekday == 7 ~ 2),
         weekday = factor(weekday, levels = c(1:2), labels = c("Weekday", "Weekend"))) %>%
  group_by(weekday, interval) %>% 
  summarise(mean = mean(steps)) %>% 
  ggplot(aes(interval, mean)) +
  geom_line() +
  facet_grid(row = vars(weekday))
```























